---
title: "Exploring Louvain Communities in Drugnet"
author: "Sean Ryan"
date: "September 28, 2026"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: false
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.align = "center", fig.width = 11, fig.height = 8,
  dpi = 150
)
```

This is a **separate exploratory page**, not part of the Assignment 2 submission. It uses the same Drugnet data and the same undirected main-component convention as the current report, then asks a new question:

> Does the network naturally separate into communities when we optimize modularity?

The **Louvain algorithm** starts with small groups and repeatedly moves actors between neighboring groups when doing so increases modularity. In plain language, modularity rewards a division where actors have more ties **inside their community** than we would expect from the network's overall pattern of connections.

The shaded polygons below do **not** determine the communities. Louvain determines membership from the edges first; the polygons only make those memberships visible.

## 1. Rebuild the same main component

```{r load-data}
library(igraph)
library(ggraph)
library(ggforce)
library(knitr)

load("drugnet.rda")
drugnet <- upgrade_graph(drugnet)

# Match the main Assignment 2 analysis:
# 1. Find the largest weakly connected component.
# 2. Collapse direction into one simple undirected relationship.
comps <- components(drugnet, mode = "weak")
largest <- which.max(comps$csize)

main_directed <- induced_subgraph(
  drugnet,
  which(comps$membership == largest)
)

main <- simplify(
  as_undirected(main_directed, mode = "collapse"),
  remove.multiple = TRUE,
  remove.loops = TRUE
)

stopifnot(is_connected(main), is_simple(main), !is_directed(main))

kable(
  data.frame(
    Actors = vcount(main),
    Relationships = ecount(main)
  ),
  caption = "Network used for Louvain community detection"
)
```

The analysis is therefore based on the same underlying social relationships used for the primary centrality calculations in Assignment 2.

## 2. Detect Louvain communities

```{r louvain}
set.seed(295)
louvain <- cluster_louvain(main)

V(main)$Community <- factor(membership(louvain))

community_summary <- data.frame(
  Community = seq_along(sizes(louvain)),
  Actors = as.integer(sizes(louvain))
)

community_summary <- community_summary[
  order(-community_summary$Actors),
]

kable(
  community_summary,
  row.names = FALSE,
  caption = "Louvain community sizes"
)

kable(
  data.frame(
    Communities = length(louvain),
    Modularity = modularity(louvain)
  ),
  digits = 4,
  caption = "Louvain result"
)
```

Louvain identifies **`r length(louvain)` communities** in the main component. The resulting modularity is **`r round(modularity(louvain), 3)`**.

A higher modularity means the observed grouping contains more within-community relationships than would be expected under the modularity null model. The number is useful for comparing alternative partitions of this same network; it is not a percentage of relationships that are "correct."

## 3. Show the communities as polygons

The node positions below use the same **stress layout** convention as the Assignment 2 report. Louvain does not use these x/y positions when finding communities.

```{r community-polygons}
set.seed(295)
main_layout <- create_layout(main, layout = "stress")
main_layout$Community <- V(main)$Community

ggraph(main_layout) +
  ggforce::geom_mark_hull(
    aes(
      x = x,
      y = y,
      group = Community,
      fill = Community,
      label = paste("Community", Community)
    ),
    alpha = 0.12,
    colour = NA,
    expand = grid::unit(3, "mm"),
    label.fill = "white",
    label.fontsize = 9,
    show.legend = FALSE
  ) +
  geom_edge_link(
    colour = "grey72",
    alpha = 0.6,
    width = 0.45
  ) +
  geom_node_point(
    aes(colour = Community),
    size = 2.6,
    alpha = 0.95
  ) +
  labs(
    title = "Louvain communities in the Drugnet main component",
    subtitle = "Community membership is calculated from network ties; shaded hulls only visualize the result",
    colour = "Community"
  ) +
  theme_graph(base_family = "sans") +
  theme(legend.position = "bottom")
```

If the polygons line up with visibly dense clusters, that is the visual consequence of the same pattern Louvain is quantifying: relatively many ties inside a community and relatively fewer ties crossing between communities.

## 4. Which actors connect the communities?

A useful next step is to compare Louvain membership with **betweenness centrality**. An actor can have high betweenness when many shortest paths between other actors pass through that position. If a high-betweenness actor sits near the boundary between Louvain communities, the graph gives an intuitive picture of potential brokerage.

```{r brokerage}
bet <- betweenness(
  main,
  directed = FALSE,
  normalized = TRUE,
  weights = NA
)

V(main)$Betweenness <- bet

top_n <- 10
top_ids <- order(bet, decreasing = TRUE)[seq_len(top_n)]
V(main)$TopBroker <- ifelse(
  seq_len(vcount(main)) %in% top_ids,
  V(main)$name,
  ""
)

main_layout$Betweenness <- V(main)$Betweenness
main_layout$TopBroker <- V(main)$TopBroker

ggraph(main_layout) +
  ggforce::geom_mark_hull(
    aes(
      x = x,
      y = y,
      group = Community,
      fill = Community
    ),
    alpha = 0.10,
    colour = NA,
    expand = grid::unit(3, "mm"),
    show.legend = FALSE
  ) +
  geom_edge_link(
    colour = "grey78",
    alpha = 0.55,
    width = 0.4
  ) +
  geom_node_point(
    aes(colour = Community, size = Betweenness),
    alpha = 0.9
  ) +
  geom_node_label(
    aes(filter = TopBroker != "", label = TopBroker),
    repel = TRUE,
    size = 3,
    fill = "white",
    label.size = 0.15
  ) +
  scale_size_continuous(
    range = c(1.8, 10),
    name = "Betweenness"
  ) +
  labs(
    title = "Louvain communities and potential brokers",
    subtitle = "Larger nodes have higher betweenness; the ten highest-betweenness actors are labeled",
    colour = "Community"
  ) +
  theme_graph(base_family = "sans") +
  theme(legend.position = "bottom")
```

This does **not** establish that a labeled actor actually controls drug distribution or information flow. The recorded edges are reported social relationships. The figure only compares two structural properties of the observed network: modularity-based community membership and shortest-path brokerage.

## 5. Count relationships inside and between communities

We can also inspect what Louvain is responding to directly in the edge data.

```{r inside-between}
edge_ends <- ends(main, E(main), names = FALSE)

same_community <-
  V(main)$Community[edge_ends[, 1]] ==
  V(main)$Community[edge_ends[, 2]]

relationship_summary <- data.frame(
  Relationship_type = c(
    "Within the same Louvain community",
    "Between different Louvain communities"
  ),
  Relationships = c(
    sum(same_community),
    sum(!same_community)
  )
)

relationship_summary$Percent <- round(
  100 * relationship_summary$Relationships / ecount(main),
  1
)

kable(
  relationship_summary,
  row.names = FALSE,
  caption = "Observed relationships within and between Louvain communities"
)
```

This table is the simplest data-level view of the pattern shown by the polygons: each edge is checked to see whether its two endpoints received the same Louvain community label.

## Interpretation

The useful idea here is not that Louvain discovers one objectively true set of social groups. Community detection depends on the network definition and the algorithm. Instead, Louvain provides a reproducible answer to:

> **What grouping produces a relatively strong concentration of relationships within groups compared with the modularity null model?**

That makes it useful alongside the Assignment 2 measures. Centrality describes **positions of individual actors**, while Louvain describes **larger structural groupings** in the network.

## Reproducibility

```{r session-info}
sessionInfo()
```
